---
title: "ClinicalAnalisys_DLBCL"
author: "Epitácio Farias, Beatriz Stransky, Patrick Terrematte"
date: "14/08/2020"
output: 
  html_document: 
    default
  github_document: 
    df_print: paged
    html_preview: FALSE
    keep_html: TRUE
  pdf_document:
    latex_engine: xelatex
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_format = "all") })
---

```{r, error=TRUE, message=FALSE, warning=FALSE, purl=FALSE, results='hide'}
## This chunk automatically generates a text .R version of this script when running within knitr.
input  = knitr::current_input()  # filename of input document
output = paste(tools::file_path_sans_ext(input), 'R', sep = '.')
knitr::purl(input,output,documentation=2,quiet=T)
# Avoid duplicate label error of knitr::purl
options(knitr.duplicate.label = 'allow')
# Code to browse the markdown file with renderized images.
knitr::opts_chunk$set(
  fig.path = "figs/render-"
)
```


```{r message=FALSE, warning=FALSE, echo = FALSE}
# Set the packages of interest
packages = c("tidyverse","skimr","finalfit","ggcharts")

# if a package is installed, it will be loaded
# otherwise, the missing package(s) will be installed and loaded
package.check <- lapply(packages, FUN = function(x) {
  if (!require(x, character.only = TRUE)) {
    install.packages(x, dependencies = TRUE)
    library(x, character.only = TRUE)
  }
})

suppressMessages(library("tidyverse"))
rm(packages)
setwd(".")
```

## 1. Data importing and visualizing

```{r message=FALSE, warning=FALSE, paged.print=TRUE}
dlbcl_clin_raw <- read_delim("DBLC_clinical/clinical.tsv", "\t", 
                            escape_double = FALSE, 
                            trim_ws = TRUE, na = c("\'--","NA"))
```

```{r echo=FALSE, message=FALSE, results='hide', paged.print=TRUE}
class(dlbcl_clin_raw) 
dim(dlbcl_clin_raw) 
names(dlbcl_clin_raw) 
glimpse(dlbcl_clin_raw)
skim(dlbcl_clin_raw) 
#View(dlbcl_clin_raw)
```

## 2. Cleaning data

Select variables based on NA count (> 50% complete is a good choice!).
<!-- # TO DO @PATRICK: simplify code NA_sum? -->
<!-- # kirc_clean <- kirc_clin_raw %>% -->
<!-- #     summarise_all(~ sum(is.na(.)))  -->

```{r}
NA_fifty <- dim(dlbcl_clin_raw)[1]/2

NA_sum <- colSums(is.na(dlbcl_clin_raw))
NA_sum <- as.data.frame(NA_sum)
NA_sum <- tibble::rownames_to_column(NA_sum, "variables")
NA_sum <- NA_sum %>%
     filter(NA_sum < NA_fifty)

dlbcl_clean <- dlbcl_clin_raw %>%
     select(one_of(NA_sum$variables))
```

Remove duplicate observations:

```{r}
dlbcl_clean0 <- dlbcl_clean %>%
     distinct_at('case_submitter_id', .keep_all = TRUE)
```

Remove nuneric variables with unique observations:  
<!-- # TO DO @PATRICK: function to select variables with unique observations? 
<!-- # kirc_cleanX <- kirc_clean1 %>% -->
<!-- #     summarise_if(is.numeric, ~ n=unique(.)) -->

```{r message=FALSE, warning=FALSE, paged.print=TRUE}
dlbcl_clean0 %>%
     select_if(is.numeric) %>%
     skim()

dlbcl_clean1 <-  dlbcl_clean0  %>%
     select(!c('days_to_diagnosis'))
```

Remove character variables with unique observations:

```{r  message=FALSE, warning=FALSE, paged.print=TRUE}
dlbcl_clean1 %>%
  select_if(is.character) %>%
  skim()

dlbcl_clean2 <- dlbcl_clean1  %>%
    select(!c('project_id', 'classification_of_tumor', 'morphology', 'progression_or_recurrence',
              'synchronous_malignancy', 'tumor_grade', 'tumor_stage','last_known_disease_status'))
```

Remove character variables with similar information - check each one!

```{r}
table(dlbcl_clean2$`site_of_resection_or_biopsy`, exclude = NULL)
table(dlbcl_clean2$`tissue_or_organ_of_origin`, exclude = NULL)

dlbcl_clean3 <- dlbcl_clean2  %>%
     select(!c('site_of_resection_or_biopsy'))
```

Remove other variables not directly related to patient - check each one!

```{r}
dlbcl_clean4 <- dlbcl_clean3  %>%
     select(!c('days_to_birth','age_at_diagnosis','icd_10_code','case_id'))
```

## 3. Changing variables names

Using snake_style 

```{r}
dlbcl_clean4 <- dlbcl_clean4 %>%
     rename(patient_id = 'case_submitter_id',
            age = 'age_at_index',
            aa_b_symp = 'ann_arbor_b_symptoms',
            aa_clin_stg = 'ann_arbor_clinical_stage',
            aa_ex_inv = 'ann_arbor_extranodal_involvement')
```

## 4. Taming data

Use lubridate for dates

```{r}
dlbcl_clean4 <- dlbcl_clean4 %>%
     mutate_if(is.character, as.factor) %>%
     mutate(patient_id = as.character(patient_id),
            age = as.integer(age),
            year_of_diagnosis = as.integer(year_of_diagnosis),
            year_of_birth = as.integer(year_of_birth),
            days_to_last_follow_up = as.integer(days_to_last_follow_up))
```

## 5. Checking NA patterns 

Check distincts types of NAs: MCAR, MAR, MNAR

```{r}
dlbcl_clean4  %>%
     missing_plot()

missing_glimpse(dlbcl_clean4)
```

## 6. Checking numeric variables

Check data distribution, plausible ranges, outliers;
Thinking about deleting outliers from dataset? Need to evaluate carefully each one!

```{r}
dlbcl_clean4 %>%
     select_if(is.numeric) %>%
     summary()
```

```{r}
ggplot(dlbcl_clean4, aes(age)) +
     geom_histogram(bins = 20, alpha = 0.8, color = "red")
```

```{r}
ggplot(dlbcl_clean4, aes(year_of_diagnosis)) +
     geom_histogram(bins = 20, alpha = 0.8, color = "red")
```

```{r}
ggplot(dlbcl_clean4, aes(days_to_last_follow_up)) +
     geom_histogram(bins = 20, alpha = 0.8, color = "red")
```

```{r}
ggplot(dlbcl_clean4, aes(year_of_birth)) +
     geom_histogram(bins = 20, alpha = 0.8, color = "red")
```

```{r}
ggplot(dlbcl_clean4, aes(x ='', y=age)) +
     geom_boxplot(width = .5) +
     geom_jitter(width = 0.05, alpha = 0.2, color = "orange")
boxplot.stats(dlbcl_clean4$age)
```

```{r}
ggplot(dlbcl_clean4) +
     geom_bar(aes(x = vital_status), color = "black", fill = "orange")
```

```{r}
ggplot(dlbcl_clean4) +
     geom_bar(aes(x = ethnicity), color = "blue", fill = "purple" )
```

```{r}
dlbcl_clean4 %>%
  filter(race %in% c('asian', 'white', 'black or african american')) %>%
  bar_chart(x = vital_status, facet = race,)

```

## 7. Checking categorical variables

Check frequency, lables and levels 

```{r}
dlbcl_clean4 %>%
     select_if(is.factor) %>%
     summary() 

# changing level names
dlbcl_clin <- dlbcl_clean4 %>%
     mutate(aa_clin_stg= fct_recode(aa_clin_stg, 'STG1'='Stage I', 'STG2'='Stage II', 'STG3'='Stage III',
                                    'STG4'='Stage IV'),
            treatment_type= fct_recode(treatment_type, 'Pharmaceutical' = 'Pharmaceutical Therapy, NOS', 
            'Radiation'='Radiation Therapy, NOS' ))

dlbcl_clin %>%
     select_if(is.factor) %>%
     summary()
```

## 8. Saving dataset

```{r}
write_csv(dlbcl_clin, path = "DBLC_clinical/dlbcl_clin.csv")

rm(dlbcl_clean4, dlbcl_clean3, dlbcl_clean2, dlbcl_clean1, dlbcl_clean0, dlbcl_clean, NA_sum, NA_fifty)
```

## Further analysis

- [Comparison and Hyphotesis test](2.correlation.md) 
- [Logistic Regression Model](3.logistic_regression.md)

```{r}
sessionInfo()
```


